Characterizing edges in signed and vector-valued graphs

### 2017-12-18 cc/cc_state_of_art.tex
fix cc chapter layout
pass over CC

As we saw, solving \pcc{} on a training set would give us a principled way of predicting edge signs.
In this section, we thus look at existing methods for \pcc{}. We start with exact methods and fixed
parameter algorithms on complete graphs, noting they implicitly assume our learning bias. However,
in practice, these methods do not scale well, especially if their assumptions are not met. Indeed,
we next present hardness results about \pcc{}, namely that it is \NPh{} to approximate with some
constant factor greater than one, and then describe some proposed approximations. Whereas this
sounds like bad news for us, we show that heuristics can solve large instances with satisfactory
results. We conclude by further highlighting the learning aspect of \pcc{} through approaches in
the active and online settings.

@@

Because of the complexity of the \pcc{}, one has to rely on approximations to solve large instances
of the problem. However, we can imagine offline signed social networks with only few nodes, in which
case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
to evaluate in practice the quality of heuristic methods, albeit in non asymptotic settings. The
methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
fixed parameter algorithms we describe now only handle complete graphs, while the next set of
methods also work on general graphs.

@@

Let us first describe the cluster editing problem, where one is given a general unsigned input graph
$H=(V,E)$ and wants to
find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
union of cliques. It is possible to alternate between an instance $H$ of cluster editing and an
instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
are deleted from $H$. To the best of our knowledge, the problem was first introduced under this
name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
whereas we shall only give important and more recent pointers here.} They show the problem is
\NPc{}, even if the number of clusters $K\geq2$ is set beforehand\footnote{The reduction is from
3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
$0.878$-approximation in the weighted $K=2$ case using the celebrated Goemans-Williamson SDP
relaxation~\autocite{MaxCutSDP95}.

@@

For that we follow~\autocite{Demaine2006} but omit the full proofs to avoid redundancy. Recall that
given a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, in \mmc{} we want to find the lightest set of edges $T\in E_H$
whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
polynomial transformation from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
are simulated in the following way: include every source and target in a clique of $n$ nodes and
connect that clique to the corresponding source or target by $n$ negative edges.

\begin{table}[b]
   \centering
   \small
   \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
   \begin{tabulary}{187mm}{lCCCC}
      \toprule
               & \multicolumn{2}{c}{\mind{}}   & \multicolumn{2}{c}{\maxa{}}                                   \\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      graph    & weighted                      & unweighted                                                     & weighted                                                     & unweighted                   \\
      \midrule
      Complete & \APXh{}~\autocite{Charikar2003} & \NPc{}~\autocite{Bansal2002}, \APXh{}~\autocite{Charikar2003}  &                                                              & \NPc{}~\autocite{Bansal2002} \\
      General  & \APXh{}~\autocites{Charikar2003}{Demaine2003} &  \APXh{}~\autocites{Charikar2003}{Emanuel2003}   & \multicolumn{2}{c}{\APXh{}~\autocite[Thm. 9]{Charikar2003}} \\
      \bottomrule
   \end{tabulary}
\end{table}

We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
4.4]{Demaine2006}) that it takes a linear time to construct a
multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
%\Todo{give the intuition of the proof}

@@

If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
within a factor $c$~\autocite{MulticutHardness06}. On general graphs, the best known approximation factor in $O(\log
k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
$O(\log|E^-|)=O(\log n^2)=O(\log n)$. This approximation ratio is achieved by first solving the
relaxation of the integer problem \eqref{eq:mindIP}, where the constraint \eqref{eq:mindInteger} is
replaced by $x_{uv}\in [0,1]$. From now on, we refer to that relaxation as the \emph{canonical
\mind{} LP}. Once this canonical \mind{} LP is solved, we interpret $x_{uv}$ as a distance: the
larger it is and the more we want $u$ and $v$ to be in different clusters. We can then use the
\regionGrow{} method~\autocite{RegionGrowing93}. The idea is to pick a random center $u$ and to add
to $u$'s cluster all the nodes at distance less than $r$ from $u$ before removing that cluster and
repeating the process. $r<\shalf$ is chosen adaptively such that the weight of positive edges
leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
inside the cluster.
%\Todo{The full proof can be found in pages 7--9 of
%\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
%doesn't have much to do with machine learning.}

However, \textcite[Theorem 2]{Charikar2003} note that the LP formulation has a poor integrality gap
when it comes to \maxa{}, thus they turn to a Semi Definite Program. Say that each cluster is
associated with a basis vector, then for each node $u$ in a cluster, we set $a_u$ to be the
corresponding basis vector. If $u$ and $v$ are in the same cluster, we then have $a_u\cdot a_v = 1$
while if they belong to different clusters, $a_u\cdot a_v = 0$. The weighted number of agreements
can then be represented by
\begin{align}
   \label{eq:maxaSDP}
   \text{maximize } & \sum_{(u,v)\in E^+} w_{uv}(a_u\cdot a_v) + \sum_{(u,v)\in E^-} w_{uv}(1-a_u\cdot a_v) \\
   \text{subject to}& \quad a_u\cdot a_u=1 \nonumber\\
   \phantom{subject to}& \quad a_u\cdot a_v\geq 0  \nonumber
\end{align}
After solving the SDP, a clustering can be obtained by a general rounding technique $H_t$: pick $t$
random hyperplanes and divides the nodes in $2^t$ clusters. \Textcite[Theorem 3]{Charikar2003} prove
that taking the best results of $H_2$ and $H_3$ gives in a $0.7664$ approximation on general graph.
This was slightly improved to $0.7666$ by \textcite{Swamy2004} with a different rounding: pick $k$
random unit vectors (called \emph{spokes}) and assign each $a_u$ to the closest spoke.

Combining \mind{} and \maxa{}, \textcite[Section 4]{Charikar2004} give a $\Omega(\frac{1}{\log n})$
approximation of the \textsc{MaxCorr} problem, which is maximizing \eqref{eq:maxa} - \eqref{eq:mind}
and can be formulated as a quadratic programming problem solved in polynomial time.

@@

\begin{table}
   \caption{Best current results on \pcc{} problems. The \enquote{easiest} setting is \maxa{} on
      complete graphs, for it admits PTASs. All others cases are \APXh{}. However, we see that on
      the diagonal (that is \mind{} on complete graphs and \maxa{} on general graphs), there exists
      constant factor approximations. This is not the case for the most \enquote{difficult} problem,
      \mind{} on general graphs. \label{tab:cc_approx}}
   \begin{tabulary}{\textwidth}{lLL}
      \toprule
      $G$      &  \mind{}                                                  & \maxa{}                                 \\
      \midrule
      Complete &  $2.06$ (and $1.5$ if $G$ is weighted and weights obey the triangular inequality) \autocite{Chawla2014}
               & PTAS from \textcite{Bansal2002} and by setting $k=\Omega(1/\epsilon)$ in \autocite{Giotis2006}      \\
      General  &  $O(\log n)$ \autocite{Charikar2003}                      & $0.7666$ \autocite{Swamy2004}           \\
      \bottomrule
   \end{tabulary}
\end{table}

@@

This concludes the presentation of the known approximation results on \pcc{}, that we gather in
\autoref{tab:cc_approx}.
To summarize, except for maximizing the number of agreements on a complete graphs, we cannot hope
for exact solutions for the other problems in the worst case, since they are \APXh{}. Moreover, even
constant factor approximation are ruled out for \mind{} on general graphs, as showed by the
equivalence with \mmc{}.
Currently there are three main approaches to approximate \mind{}. On a general graph,
one first solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
resulting in an $O(\log n)$ approximation. On complete graphs, the simplest solution is the
\ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
the proof of \autoref{thm:weakbal-consistent}. The approximation factor of this general idea can be
improved by interpreting the fractional solution of the linear program as probability to be included
in the pivot's cluster. The cost of solving linear programs with so many variables and constraints
being prohibitive, real world general graphs requires to move from methods with guarantees to more
heuristic approaches we describe next. Before that, we make two remarks. The first one regards a
line of work improving scalability of these approximations by presenting distributed solutions. The
second one is concerned with the case where the number of clusters is set beforehand, and proves
that the problem becomes easier.

@@

\Textcite{Bonchi2013} suggest a different paradigm to solve \pcc{}, that can be applied in a
distributed setting to obtain a scalable approach. Namely, given a node $u$, they want to output a
globally consistent cluster index $\cluster(u)$ while making at most $t$ queries to a sign oracle.
Here $t$ is a parameter that depends on the quality of clustering produced but not on the size of
the graph. And because this procedure is local to each node, it can be run in parallel. Finally, one
can get a full clustering by computing $\cluster(u)$ for all the nodes in the graph. Despite the
problem being apparently more challenging, they obtain approximation factors that are close to the
best known (which in this model would make $\Omega(n^2)$ total queries). More precisely, they use
two techniques.
The first one is inspired by \ccpivot{}. It starts by finding a good set of pivots, seeing the
problem as finding a maximal independent set on a sampled part of the positive graph. Then, for a
given node, it finds the closest such pivot or creates a singleton. Given a quality parameter
$\epsilon\in(0,1)$, it yields a $4\cdot OPT + \epsilon n^2$ approximation of \mind{} requiring
$O(\frac{n}{\epsilon})$ time and queries~\autocite[Theorem 3.3]{Bonchi2013}.
Roughly stated, the second one relies on an existing low-rank approximation of the adjacency matrix,
that partition the graph into similar sized classes such that edges between those classes behave as
in a random graph. This initial partition is \enquote{coarsened} into a good clustering by
considering all possible ways of assigning classes to clusters. It gives an $OPT + \epsilon n^2$
additive approximation  for \mind{} that runs in time $n \cdot poly(\nicefrac{1}{\epsilon}) +
2^{poly(\nicefrac{1}{\epsilon})}$ \autocite[Corollary 3.7]{Bonchi2013}.
\end{aside}

\begin{aside}
As mentioned earlier, not having to set the number of clusters is an
attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
problem was studied by \textcite{Giotis2006} on general graphs and we compile their results in
\autoref{tab:cc_fixed}. On complete unweighted graphs and with $K$ being the number of clusters, they
provide PTASs for \maxa{}$[K]$ running in $nk^{O(\epsilon^{-3}\log(\frac{K}{\epsilon}))}$ time and for
\mind{}$[K]$ running in $n^{O\left(\epsilon^{-2} 9^K\right)}\log(n)$ time. The latter was improved by
\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}K^6\log d\right)}$
and that can handle weighted graphs. For $K=2$, \mind$[2]$ admits a faster local search method with a
factor $2$ approximation~\autocite{Coleman2008}.
For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
\maxa{}$[K]$ under the condition that the ratio between the largest and smallest weights is bounded by a
constant.
\begin{center}
   \captionsetup{font=small}
   \captionof{table}{\label{tab:cc_fixed} Approximation results for \pcc{} on general graphs with $K$
   clusters}
   % 187mm - 3em = https://tex.stackexchange.com/a/8263
   \begin{tabulary}{177mm}{lLL}
      \toprule
      $K$	 & $2$ & $\geq 3$ \\
      \midrule
      \maxa{} & $0.878$ (improved to $0.884$ by \autocite{Mitra2009}) & $0.7666$
      \autocite{Swamy2004} (and it can be proved that this creates at most $6$ clusters) \\
      \mind{} & $O(\sqrt{\log n})$ as it reduces to Min 2CNF Deletion for which \textcite{min2CNF05}
      give such an approximation &
      this can be reduced from $K$-coloring, which for any $\epsilon > 0$ is \NPc{} to approximate
      within $n^{1-\epsilon}$ \autocite{InnaproxChroma07} \\
      \bottomrule
   \end{tabulary}
\end{center}
\end{aside}

@@

After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
\textcite{Elsner2009} describe three folklore heuristics that start with empty clusters and add node
one by one: \enquote{The \textsc{Best} algorithm adds each node $u$ to the cluster with the
strongest $w^\pm$ connecting to $u$, or to a new singleton if none of the $w^\pm$ are positive.
The \textsc{First} algorithm adds each node $u$ to the cluster containing the most recently
considered node $v$ with $w^\pm_{uv} > 0$. The \textsc{Vote} algorithm adds each node to the cluster
that minimizes the \pcc{} objective, \ie{} to the cluster maximizing the total net weight
or to a singleton if no total is positive.} Empirically, \textsc{Vote} turns out to be the best.
Among other related heuristics, \textcite{mergingHeuristics14} describe the random maximum merging
algorithm, that starts with singleton clusters and keep merging two clusters chosen at random among
those whose merge would result in the maximum improvement of the score function. This runs in
$O(n^2\log n)$, and empirically results in fewer disagreements than \ccpivot{}.
Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
greedily build a clustering in fashion similar to \textsc{Vote}, albeit with more randomness in the
node ordering. This clustering is locally improved by moving blocks of $r\in\{1,2\}$ nodes from one
cluster to another as long as the number of disagreements decreases, a phase called neighborhood
descent. The algorithm then enters an inner loop where the current clustering is perturbed by $t$
random one-node-moves and updated if a subsequent neighborhood descent can improve it compared with
before the perturbation. The authors note that both the outer loop and the neighborhood descent can
be run in parallel, which allow them to process a \np{10000} nodes graph on 10 cores in around 700
seconds. \Textcite{heuristicCE16} show how to use ILS to generate initial solutions of the Cluster
Editing problem that are then fed to an integer program.
% This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
% approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!}
Instead of
moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
graph, $G^+$ is a disjoint union of cliques in which all the nodes of a given cluster share the shame
neighbors. Their algorithm is initialized with the set $E_s^+$ of edges where $w^+_{uv} > w^-_{uv}$,
and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the
canonical LP to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
$L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
otherwise~\autocite{AltOptimLP13}. Indeed, $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
is the cluster assignment. By relaxing the integer constraints on $L$, it is possible to do
alternate optimization on $L$ and $\cluster$.

>>>>

As we saw, solving \pcc{} on a training set would give us a principled way of predicting edge signs.
In this section, we thus look at existing methods for \pcc{}. We start with exact methods and fixed
parameter algorithms on complete graphs, noting they implicitly assume our learning bias. However,
in practice, these methods do not scale well, especially if their assumptions are not met. Indeed,
we next present hardness results about \pcc{}, namely that it is \NPh{} to approximate with some
constant factor greater than one, and then describe some proposed approximations. Although this
appears rather negative at first, we show that heuristics can solve large instances with satisfactory
results. We conclude by further highlighting the learning aspect of \pcc{} through approaches in
the active and online settings.

@@

Because of the complexity of the \pcc{}, one has to rely on approximations to solve large instances
of the problem. However, we can imagine offline signed social networks with only few nodes, in which
case it is reasonable to expect to find the optimal clustering. Furthermore, this can also be useful
to evaluate in practice the quality of heuristic methods, albeit in non asymptotic settings. The
methods used to solve \pcc{} exactly depend on the type of graph considered. More precisely, the
fixed parameter algorithms we describe now only handle complete graphs, while the next set of
methods also work on general graphs.

@@

Let us first describe the cluster editing problem, where one is given a general unsigned input graph
$H=(V,E)$ and wants to
find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
union of cliques. It is possible to alternate between an instance $H$ of cluster editing and an
instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
be a complete signed graph where $V$ is the nodeset of $h$, $E^+$ is the set $E$ of edges of $H$ and
$E^-$ is the set of all
edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
are deleted from $H$. To the best of our knowledge, the problem was first introduced under this
name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
whereas we shall only give important and more recent pointers here.} They show the problem is
\NPc{}, even if the number of clusters $K\geq2$ is set beforehand\footnote{The reduction is from
3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
$0.878$-approximation in the weighted $K=2$ case using the celebrated Goemans-Williamson SDP
relaxation~\autocite{MaxCutSDP95}.

@@

For that we follow~\autocite{Demaine2006} but omit the full proofs to avoid redundancy. In \mmc{},
given a weighted graph $H=(V_H, E_H, w_H)$ and a collection of $k$ sources and targets
$S=\{(s_i,t_i)\}_{i=1}^k \in V_H^2$, we want to find the lightest set of edges $T\in E_H$
whose removal disconnects every pair $s_i$ from the corresponding $t_i$. Let us first describe a
polynomial transformation from \mmc{} to \pcc{}. Namely, given an instance $(H, S)$ of \mmc{}, we let
$W_H=\sum_{e\in E_H} w_H(e)$ be the total weight of $H$ and $G_H$ be the same graph as $H$ with all
its edges labeled positively. We then connect every pair $(s_i, t_i)$ by a negative edge of weight
$W_H+1$. One can check (\autocite[Theorem 4.7]{Demaine2006}) that a multicut $T$ in $H$ with weight
$W$ induces a clustering in $G_H$ of weighted disagreement $W$ by the connected components of
$G^+\setminus T$. Likewise, a clustering of $G_H$ with weight $W$ induces a multicut on $H$ with
weight as most $W$. In the unweighted case, the reduction is similar but the heavy negative edges
are simulated in the following way: include every source and target in a clique of $n$ nodes and
connect that clique to the corresponding source or target by $n$ negative edges.

\begin{table}[tb]
\begin{adjustwidth}{-2cm}{}
   \centering
   \small
   \caption{Hardness results of \pcc{}} \label{tab:cc_cpx}
   \begin{tabulary}{179mm}{lCCCC}
      \toprule
               & \multicolumn{2}{c}{\mind{}}   & \multicolumn{2}{c}{\maxa{}}                                   \\
      \cmidrule(r){2-3}
      \cmidrule(r){4-5}
      graph    & weighted                      & unweighted                                                     & weighted                                                     & unweighted                   \\
      \midrule
      Complete & \APXh{}~\autocite{Charikar2003} & \NPc{}~\autocite{Bansal2002}, \APXh{}~\autocite{Charikar2003}  &                                                              & \NPc{}~\autocite{Bansal2002} \\
      General  & \APXh{}~\autocites{Charikar2003}{Demaine2003} &  \APXh{}~\autocites{Charikar2003}{Emanuel2003}   & \multicolumn{2}{c}{\APXh{}~\autocite[Thm. 9]{Charikar2003}} \\
      \bottomrule
   \end{tabulary}
\end{adjustwidth}
\end{table}

We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v) \in E^-$ of
weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
4.4]{Demaine2006}) that it takes a linear time to construct a
multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
%\Todo{give the intuition of the proof}

@@

If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$, it is \NPh{} to approximate \mmc{}
within a factor $c$~\autocite{MulticutHardness06}. On general graphs, the best known approximation factor in $O(\log
k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
$O(\log|E^-|)=O(\log n^2)=O(\log n)$. This approximation ratio is achieved by first solving the
relaxation of the integer problem \eqref{eq:mindIP}, where the constraint \eqref{eq:mindInteger} is
replaced by $x_{uv}\in [0,1]$. From now on, we refer to that relaxation as the \emph{canonical
\mind{} LP}. Once this canonical \mind{} LP is solved, we interpret $x_{uv}$ as a distance: the
larger it is and the more we want $u$ and $v$ to be in different clusters. We can then use the
\regionGrow{} method~\autocite{RegionGrowing93}. The idea is to pick a random center $u$ and to add
to $u$'s cluster all the nodes at distance less than $r$ from $u$ before removing that cluster and
repeating the process. $r<\shalf$ is chosen adaptively such that the weight of positive edges
leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
inside the cluster.
%\Todo{The full proof can be found in pages 7--9 of
%\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
%doesn't have much to do with machine learning.}

However, \textcite[Theorem 2]{Charikar2003} note that the LP formulation has a poor integrality gap
when it comes to \maxa{}, thus they turn to a Semi Definite Program. Say that each cluster is
associated with a basis vector, then for each node $u$ in a cluster, we set $a_u$ to be the
corresponding basis vector. If $u$ and $v$ are in the same cluster, we then have $a_u\cdot a_v = 1$
while if they belong to different clusters, $a_u\cdot a_v = 0$. The weighted number of agreements
can then be represented by
\begin{align}
   \label{eq:maxaSDP}
   \text{maximize } & \sum_{(u,v)\in E^+} w_{uv}(a_u\cdot a_v) + \sum_{(u,v)\in E^-} w_{uv}(1-a_u\cdot a_v) \\
   \text{subject to}& \quad a_u\cdot a_u=1 \nonumber\\
   \phantom{subject to}& \quad a_u\cdot a_v\geq 0  \nonumber
\end{align}
After solving the SDP, a clustering can be obtained by a general rounding technique $H_t$: pick $t$
random hyperplanes and divide the nodes in $2^t$ clusters. \Textcite[Theorem 3]{Charikar2003} prove
that taking the best results of $H_2$ and $H_3$ gives in a $0.7664$ approximation on general graph.
This was slightly improved to $0.7666$ by \textcite{Swamy2004} with a different rounding: pick $k$
random unit vectors (called \emph{spokes}) and assign each $a_u$ to the closest spoke.

Combining \mind{} and \maxa{}, \textcite[Section 4]{Charikar2004} give a $\Omega(\frac{1}{\log n})$
approximation of the \textsc{MaxCorr} problem, which is maximizing \eqref{eq:maxa} $-$ \eqref{eq:mind}
and can be formulated as a quadratic programming problem solved in polynomial time.

@@

\begin{table}
  \caption[Best current results on \pcc{} problems]{Best current results on \pcc{} problems. The \enquote{easiest} setting is \maxa{} on
      complete graphs, for it admits PTASs. All others cases are \APXh{}. However, we see that on
      the diagonal (that is \mind{} on complete graphs and \maxa{} on general graphs), there exists
      constant factor approximations. This is not the case for the most \enquote{difficult} problem,
      \mind{} on general graphs. \label{tab:cc_approx}}
   \begin{tabulary}{\textwidth}{lLL}
      \toprule
      $G$      &  \mind{}                                                  & \maxa{}                                 \\
      \midrule
      Complete &  $2.06$ (and $1.5$ if $G$ is weighted and weights obey the triangular inequality) \autocite{Chawla2014}
               & PTAS from \textcite{Bansal2002} and by setting $k=\Omega(1/\epsilon)$ in \autocite{Giotis2006}      \\
      General  &  $O(\log n)$ \autocite{Charikar2003}                      & $0.7666$ \autocite{Swamy2004}           \\
      \bottomrule
   \end{tabulary}
\end{table}

@@

This concludes the presentation of the known approximation results on \pcc{}, that we gather in
\autoref{tab:cc_approx}.
To summarize, except for maximizing the number of agreements on a complete graphs, we cannot hope
for exact solutions for the other problems in the worst case, since they are \APXh{}. Moreover, even
constant factor approximation are ruled out for \mind{} on general graphs, as showed by the
equivalence with \mmc{}.
Currently there are three main approaches to approximate \mind{}. First, on a general graph,
we can solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
resulting in an $O(\log n)$ approximation. Second, on complete graphs, the simplest solution is the
\ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
the proof of \autoref{thm:weakbal-consistent}. Third, the approximation factor of this general idea can be
improved by interpreting the fractional solution of the linear program as probability to be included
in the pivot's cluster. The cost of solving linear programs with so many variables and constraints
being prohibitive, real world general graphs requires to move from methods with guarantees to more
heuristic approaches we describe next. Before that, we make two remarks. The first one regards a
line of work improving scalability of these approximations by presenting distributed solutions. The
second one is concerned with the case where the number of clusters is set beforehand, and proves
that the problem becomes easier.

@@

\Textcite{Bonchi2013} suggest a different paradigm to solve \pcc{}, that can be applied in a
distributed setting to obtain a scalable approach. Namely, given a node $u$, they want to output a
globally consistent cluster index $\cluster(u)$ while making at most $t$ queries to a sign oracle.
Here $t$ is a parameter that depends on the quality of clustering produced but not on the size of
the graph. And because this procedure is local to each node, it can be run in parallel. Finally, one
can get a full clustering by computing $\cluster(u)$ for all the nodes in the graph. Despite the
problem being apparently more challenging, they obtain approximation factors that are close to the
best known (which in this model would make $\Omega(n^2)$ total queries). More precisely, they use
two techniques.
The first one is inspired by \ccpivot{}. It starts by finding a good set of pivots, seeing the
problem as finding a maximal independent set on a sampled part of the positive graph. Then, for a
given node, it finds the closest such pivot or creates a singleton. Given a quality parameter
$\epsilon\in(0,1)$, it yields a $4\cdot OPT + \epsilon n^2$ approximation of \mind{} requiring
$O(\frac{n}{\epsilon})$ time and queries~\autocite[Theorem 3.3]{Bonchi2013}.
Roughly stated, the second technique relies on an existing low-rank approximation of the adjacency matrix,
that partition the graph into similar sized classes such that edges between those classes behave as
in a random graph. This initial partition is \enquote{coarsened} into a good clustering by
considering all possible ways of assigning classes to clusters. It gives an $OPT + \epsilon n^2$
additive approximation  for \mind{} that runs in time $n \cdot poly(\nicefrac{1}{\epsilon}) +
2^{poly(\nicefrac{1}{\epsilon})}$ \autocite[Corollary 3.7]{Bonchi2013}.
\end{aside}

\begin{aside}
As mentioned earlier, not having to set the number of clusters is an
attractive feature of the \pcc{} problem, but in some cases we may want to use prior knowledge.  The
problem was studied by \textcite{Giotis2006} on general graphs and we compile their results in
\autoref{tab:cc_fixed}. On complete unweighted graphs and with $K$ being the number of clusters, they
provide PTASs for \maxa{}$[K]$ running in $nk^{O(\epsilon^{-3}\log(\frac{K}{\epsilon}))}$ time and for
\mind{}$[K]$ running in $n^{O\left(\epsilon^{-2} 9^K\right)}\log(n)$ time. The latter was improved by
\textcite{LinearMinPTAS09}, with a PTAS running in $n^2 2^{O\left(\epsilon^{-3}K^6\log d\right)}$
and that can handle weighted graphs. For $K=2$, \mind$[2]$ admits a faster local search method with a
factor $2$ approximation~\autocite{Coleman2008}.
For complete weighted graph, \textcite{WeightedMaxAPTAS08} provide a PTAS for
\maxa{}$[K]$ under the condition that the ratio between the largest and smallest weights is bounded by a
constant.
\begin{center}
   \captionsetup{font=small}
   \captionof{table}{\label{tab:cc_fixed} Approximation results for \pcc{} on general graphs with $K$
   clusters}
   % 187mm - 3em = https://tex.stackexchange.com/a/8263
   \begin{tabulary}{150mm}{lLL}
      \toprule
      $K$	 & $2$ & $\geq 3$ \\
      \midrule
      \maxa{} & $0.878$ (improved to $0.884$ by \autocite{Mitra2009}) & $0.7666$
      \autocite{Swamy2004} (and it can be proved that this creates at most $6$ clusters) \\
      \mind{} & $O(\sqrt{\log n})$ as it reduces to Min 2CNF Deletion for which \textcite{min2CNF05}
      give such an approximation &
      this can be reduced from $K$-coloring, which for any $\epsilon > 0$ is \NPc{} to approximate
      within $n^{1-\epsilon}$ \autocite{InnaproxChroma07} \\
      \bottomrule
   \end{tabulary}
\end{center}
\end{aside}

@@

After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
\textcite{Elsner2009} describe three folklore heuristics that start with empty clusters and add node
one by one: \enquote{The \textsc{Best} algorithm adds each node $u$ to the cluster with the
strongest $w^\pm$ connecting to $u$, or to a new singleton if none of the $w^\pm$ are positive.
The \textsc{First} algorithm adds each node $u$ to the cluster containing the most recently
considered node $v$ with $w^\pm_{uv} > 0$. The \textsc{Vote} algorithm adds each node to the cluster
that minimizes the \pcc{} objective, \ie{} to the cluster maximizing the total net weight
or to a singleton if no total is positive.} Empirically, \textsc{Vote} turns out to be the best.
Among other related heuristics, \textcite{mergingHeuristics14} describe the random maximum merging
algorithm, that starts with singleton clusters and keep merging two clusters chosen at random among
those whose merge would result in the maximum improvement of the score function. This runs in
$O(n^2\log n)$, and empirically results in fewer disagreements than \ccpivot{}.
Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
greedily building a clustering in a fashion similar to \textsc{Vote}, albeit with more randomness in the
node ordering. This clustering is locally improved by moving blocks of $r\in\{1,2\}$ nodes from one
cluster to another as long as the number of disagreements decreases, a phase called neighborhood
descent. The algorithm then enters an inner loop where the current clustering is perturbed by $t$
random one-node-moves and updated if a subsequent neighborhood descent can improve it compared with
before the perturbation. The authors note that both the outer loop and the neighborhood descent can
be run in parallel, which allow them to process a \np{10000} nodes graph on 10 cores in around 700
seconds. \Textcite{heuristicCE16} show how to use ILS to generate initial solutions of the Cluster
Editing problem that are then fed to an integer program.
% This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
% approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!}
Instead of
moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
graph, $G^+$ is a disjoint union of cliques in which all the nodes of a given cluster share the shame
neighbors. Their algorithm is initialized with the set $E_s^+$ of edges where $w^+_{uv} > w^-_{uv}$,
and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the
canonical LP to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
$L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
otherwise~\autocite{AltOptimLP13}. Indeed, $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
is the cluster assignment. By relaxing the integer constraints on $L$, it is possible to do
alternate optimization on $L$ and $\cluster$.
